---
title: "Disaster Relief Project Part 1 and 2"
author: "Yihnew Eshetu"
date: "2/26/2020"
output: pdf_document
---

# Disaster Relief Project Part 1
```{r, message = FALSE}
library(ISLR)
library(regclass)
library(caret)
library(MASS)
library(class)
library(pROC)
library(ROCR)
library(nnet)
library(randomForest)
library(e1071)
```

## Read in Haiti csv file
```{r}
haiti.image = read.csv("HaitiPixels.csv")
attach(haiti.image)
```

## Set seed and splitting 
```{r}
set.seed(10)
data.size = nrow(haiti.image)
split.size = data.size/10
permutation = sample(data.size)
haiti.image$Class = ifelse(haiti.image$Class == 'Blue Tarp', 'B', 'NB')
haiti.image$Class = as.factor(haiti.image$Class)
```

## KNN 10x Cross-Validation
```{r}
knn.function = function(kvalue) {
  knn.acc = 0
  for (i in 1:10) {
    haiti.image.sample = permutation[((i - 1) * split.size + 1) : (i * split.size)]
    
    haiti.image.test = as.data.frame(haiti.image[haiti.image.sample,])
    haiti.image.train = as.data.frame(haiti.image[-haiti.image.sample,])
   
    knn.pred = knn(haiti.image.train[-1], haiti.image.test[-1], cl =  haiti.image.train[,1], kvalue)
    knn.acc = knn.acc + mean(knn.pred == haiti.image.test[,1])
  }
  return((knn.acc/10)*100)
}

k.value.acc = c('1' = 0, '2' = 0, '3' = 0, 
                  '4' = 0, '5' = 0, '6' = 0, 
                  '7' = 0, '8' = 0, '9' = 0, 
                  '10' = 0, '11' = 0, '12' = 0,
                  '13' = 0, '13' = 0, '15' = 0)

start_time = Sys.time()
for (k in 1:15){
  k.value.acc[k] = knn.function(k)
}
end_time = Sys.time()
end_time - start_time

k.value.acc[which.max(k.value.acc)]

plot(seq(1,15), k.value.acc,type = 'b', xlab = 'K Value', ylab = 'Accuracy', main = paste0('K = ', which.max(k.value.acc)) )
points(which.max(k.value.acc), k.value.acc[which.max(k.value.acc)], col ='red', pch = 20)
```

## LDA 10x Cross-Validation
```{r}
lda.function = function() {
  lda.acc = 0
  for (i in 1:10) {
    haiti.image.sample = permutation[((i - 1) * split.size + 1) : (i * split.size)]
    
    haiti.image.test = as.data.frame(haiti.image[haiti.image.sample,])
    haiti.image.train = as.data.frame(haiti.image[-haiti.image.sample,])
    
    lda.fit = lda(Class ~ Red + Green + Blue, data = haiti.image.train)
    lda.pred = predict(lda.fit, haiti.image.test)
    lda.acc = lda.acc + mean(lda.pred$class == haiti.image.test[,1])
  }
  return((lda.acc/10)*100)
}
start_time = Sys.time()
lda.function()
end_time = Sys.time()
end_time - start_time
```

## QDA 10x Cross-Validation
```{r}
qda.function = function() {
  qda.acc = 0
  for (i in 1:10) {
    haiti.image.sample = permutation[((i - 1) * split.size + 1) : (i * split.size)]
    
    haiti.image.test = as.data.frame(haiti.image[haiti.image.sample,])
    haiti.image.train = as.data.frame(haiti.image[-haiti.image.sample,])
    
    qda.fit = qda(Class ~ Red + Green + Blue, data = haiti.image.train)
    qda.pred = predict(qda.fit, haiti.image.test)
    qda.acc = qda.acc + mean(qda.pred$class == haiti.image.test[,1])
  }
  return((qda.acc/10)*100)
}
start_time = Sys.time()
qda.function()
end_time = Sys.time()
end_time - start_time
```

## Logistic Regression
```{r, message = FALSE}
log.function = function() {
  log.acc = 0
  for (i in 1:10) {
    haiti.image.sample = permutation[((i - 1) * split.size + 1) : (i * split.size)]
    
    haiti.image$Class = factor(haiti.image$Class)
    levels(haiti.image$Class)
    haiti.image.test = haiti.image[haiti.image.sample,]
    haiti.image.train = haiti.image[-haiti.image.sample,]
    
    log.fit = multinom(Class ~ Red + Green + Blue, data = haiti.image.train, family = binomial)
    log.pred = predict(log.fit, type = 'class', haiti.image.test)
    log.acc = log.acc + mean(log.pred == haiti.image.test[,1])
  }
  return((log.acc/10)*100)
}
start_time = Sys.time()
log.function()
end_time = Sys.time()
end_time - start_time
```

# Disaster Relief Project Part 2

## Random Forest 10x Cross-Validation
### Pick the best m and ntree value
```{r}
er.rf = NA
ntree.rf = NA
for (i in 1:3){
  set.seed(10)
  train.set = sample(nrow(haiti.image), size = nrow(haiti.image)*.6)
  haiti.image.train = haiti.image[train.set,]
  haiti.image.test = haiti.image[-train.set,]
  
  for (j in c(100, 200, 300, 400, 500)){
    low.er = 100
    low.ntree = 0
    
    haiti.image.rf = randomForest(Class ~ Red + Green + Blue, data = haiti.image.train,
                             mtry = i, ntree = j, importance = TRUE)
  
    yhat.rf =  predict(haiti.image.rf, newdata = haiti.image.test)
    er = 1 - mean(yhat.rf == haiti.image.test$Class)
    low.er = ifelse(low.er < er, low.er, er)
    low.ntree = ifelse(low.er < er, low.ntree, j)}
   
  er.rf[i] = low.er
  ntree.rf[i] = low.ntree
}
plot(1:3, er.rf, type = 'b', xlab = 'm size')
points(which.min(er.rf), 
       er.rf[which.min(er.rf)], col = 'red', pch = 20)
```

### Random Forest 10x Cross-Validation
```{r}
rf.function = function(m, n) {
  rf.acc = 0
  for (i in 1:10) {
    haiti.image.sample = permutation[((i - 1) * split.size + 1) : (i * split.size)]
    
    haiti.image.test = as.data.frame(haiti.image[haiti.image.sample,])
    haiti.image.train = as.data.frame(haiti.image[-haiti.image.sample,])
    
    rf.fit = randomForest(Class ~ Red + Green + Blue, data = haiti.image.train, 
                          mtry = m, ntree = n, importance = TRUE)
    rf.pred = predict(rf.fit, newdata = haiti.image.test)
    rf.acc = rf.acc + mean(rf.pred == haiti.image.test[,1])
  }
  return((rf.acc/10)*100)
}
start_time = Sys.time()
which.min(er.rf)
ntree.rf[which.min(er.rf)]
rf.function(which.min(er.rf), ntree.rf[which.min(er.rf)])
end_time = Sys.time()
end_time - start_time
```

## SVM 10x Cross-Validation

### Linear
```{r}
cross = tune.control(cross = 10)
start_time = Sys.time()
haiti.image.svm.linear = tune(svm, Class ~ ., data =  haiti.image, kernel = 'linear', 
                ranges = list(cost = c(1, 10, 25, 50)),
                tunecontrol = cross)
end_time = Sys.time()
end_time - start_time

haiti.image.svm.linear$best.parameters
haiti.image.svm.linear.acc = (1 - haiti.image.svm.linear$best.performance)*100
haiti.image.svm.linear.acc
```

### Polynomial
```{r}
start_time = Sys.time()
haiti.image.svm.polynomial = tune(svm, Class ~ ., data = haiti.image, kernel = 'polynomial', 
                           ranges = list(cost = c(1, 5, 10, 50), 
                                     degree = c(2, 4, 6)),  tunecontrol = cross)
end_time = Sys.time()
end_time - start_time

haiti.image.svm.polynomial$best.parameters
haiti.image.svm.polynomial.acc = (1 - haiti.image.svm.polynomial$best.performance)*100
haiti.image.svm.polynomial.acc
```

### Radial
```{r}
start_time = Sys.time()
haiti.image.svm.radial = tune(svm, Class ~ ., data = haiti.image, kernel = 'radial', 
                       ranges = list(cost = c(11, 14, 16), 
                                     gamma = c(8, 9, 10)),  tunecontrol = cross)
end_time = Sys.time()
end_time - start_time

haiti.image.svm.radial$best.parameters
haiti.image.svm.radial.acc = (1 - haiti.image.svm.radial$best.performance)*100
haiti.image.svm.radial.acc
```

## Hold Out Data
### Preprocessing and Cleaning of csv files
```{r}
NB.57 = read.csv('orthovnir057_ROI_NON_Blue_Tarps.csv', skip = 7)[, c('B1', 'B2', 'B3')]
NB.57$Class = 'NB'

B.67 = read.csv('orthovnir067_ROI_Blue_Tarps.csv', skip = 7)[, c('B1', 'B2', 'B3')]
B.67$Class = 'B'

NB.67 = read.csv('orthovnir067_ROI_NOT_Blue_Tarps.csv', skip = 7)[, c('B1', 'B2', 'B3')]
NB.67$Class = 'NB'

B.69 = read.csv('orthovnir069_ROI_Blue_Tarps.csv', skip = 7)[, c('B1', 'B2', 'B3')]
B.69$Class = 'B'

NB.69 = read.csv('orthovnir069_ROI_NOT_Blue_Tarps.csv', skip = 7)[, c('B1', 'B2', 'B3')]
NB.69$Class = 'NB'

B.78 = read.csv('orthovnir078_ROI_Blue_Tarps.csv', skip = 7)[, c('B1', 'B2', 'B3')]
B.78$Class = 'B'

NB.78 = read.csv('orthovnir078_ROI_NON_Blue_Tarps.csv', skip = 7)[, c('B1', 'B2', 'B3')]
NB.78$Class = 'NB'
```

### Merging all csv files into dataframe
```{r}
hold.out.data = as.data.frame(rbind(NB.57, B.67, NB.67, B.69, NB.69, B.78, NB.78))
hold.out.data$Class = as.factor(hold.out.data$Class)
colnames(hold.out.data) <- c('Red', 'Green', 'Blue', 'Class')
```

## KNN on Hold Out Data
```{r}
start_time = Sys.time()
knn.pred = knn(haiti.image[-1], hold.out.data[-4], cl = haiti.image[,1], k = which.max(k.value.acc))
end_time = Sys.time()
end_time - start_time

knn.acc = mean(knn.pred == hold.out.data[,4])*100
knn.acc
```

## LDA on Hold Out Data
```{r, warning =  FALSE}
start_time = Sys.time()
lda.fit = lda(Class ~ ., data = haiti.image)
lda.pred = predict(lda.fit, hold.out.data)
end_time = Sys.time()
end_time - start_time

lda.acc = mean(lda.pred$class == hold.out.data[,4])*100
lda.acc
```

## QDA on Hold Out Data
```{r, warning =  FALSE}
start_time = Sys.time()
qda.fit = qda(Class ~ ., data = haiti.image)
qda.pred = predict(qda.fit, hold.out.data)
end_time = Sys.time()
end_time - start_time

qda.acc = mean(qda.pred$class == hold.out.data[,4])*100
qda.acc
```

## Logistic Regression on Hold Out Data
```{r}
start_time = Sys.time()
log.fit = multinom(Class ~ Red + Green + Blue, data = haiti.image, family = binomial)
log.pred = predict(log.fit, type = 'class', hold.out.data)
end_time = Sys.time()
end_time - start_time

log.acc = mean(log.pred == hold.out.data[,4])*100
log.acc
```

## Random Forest on Hold Out Data
```{r}
start_time = Sys.time()
rf.fit = randomForest(Class ~ ., data = haiti.image, 
                          mtry = which.min(er.rf), ntree = ntree.rf[which.min(er.rf)], importance = TRUE)
rf.pred = predict(rf.fit, newdata = hold.out.data)
end_time = Sys.time()
end_time - start_time

rf.acc =  mean(rf.pred == hold.out.data[,4])*100
rf.acc
```

### SVM:Radial on Hold Out Data
```{r}
start_time = Sys.time()
haiti.image.svm.radial = svm(Class ~ ., data = haiti.image, kernel = 'radial', cost = 14, gamma = 9)
hold.out.svm.radial.pred = predict(haiti.image.svm.radial, newdata = hold.out.data)
end_time = Sys.time()
end_time - start_time

hold.out.svm.radial.pred.acc = mean(hold.out.svm.radial.pred == hold.out.data$Class)*100
hold.out.svm.radial.pred.acc
```

